Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-14 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

\begin{definition}[QIVPM]
  Assume a collection of intervals $\mathscr{I}$ including $\imposs$
  and $\necess$ with addition and scalar multiplication defined as
  follows:
  \begin{subequations}\label{eq:interval-operations}
  \begin{eqnarray}
   &  & [\ell_{1},r_{1}]+[\ell_{2},r_{2}]=[\ell_{1}+\ell_{2},r_{1}+r_{2}]\textrm{ and}\\
   &  & x[\ell,r]=\begin{cases}
  [x\ell,xr] & \textrm{for }x\ge0\,;\\{}
  [xr,x\ell] & \textrm{for }x\le0\,.
  \end{cases}
  \end{eqnarray}
  \end{subequations}
  Then we take a QIVPM~$\bar{\mu}$ to be an assignment of an interval to each
  event (projection operator~$P$) subject to the following constraints:
  \begin{subequations}\label{eq:QIVPM-constraints}
  \begin{eqnarray}
   &  & \bar{\mu}(\mathbb{0})=\imposs\,,\\
   &  & \bar{\mu}(\mathbb{1})=\necess\,,\\
   &  & \bar{\mu}\left(\mathbb{1}-P\right)= \necess -\bar{\mu}\left(P\right)\,,
  \end{eqnarray}
  \end{subequations}
  and satisfying for each pair of \emph{commuting} projectors~$P_0$
  and~$P_1$ with $P_0P_1=P_1P_0$,
\begin{equation}
\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,.
\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{definition}
\noindent The first three constraints,
Eqs.~(\ref{eq:QIVPM-constraints}), are the direct counterpart of the
corresponding ones for classical IVPMs.  With the understanding that
the union of classical sets $E_1\cup E_2$ is replaced by
$P_0+P_1-P_0P_1$ in the case of quantum projection
operators~\cite{Griffiths2003}, the last condition,
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}), is a
direct counterpart of the convexity condition of
Eq.~(\ref{eq:classicalconvex}). Thus our definition of QIVPMs merges
classical IVPMs with quantum probability measures.

@@

The expectation value of an observable relative to a 0-deterministic
QIVPM is fully determinate and is equal to one of the eigenvalues of
that observable. To see this, note that given an orthonormal basis
$\Omega=\{ \ket{\psi_0},\ket{\psi_1},\ldots\}$, a 0-deterministic
QIVPM must map exactly one of the projectors $\proj{\psi_i}$ to
$\necess$ and all others to $\imposs$. This is because by
Eq.~(\ref{eq:QIVPM-constraints}b), we have
$\bar{\mu}(\Sigma_i{\proj{\psi_i}})=\necess$ and by inductively applying
Eq.~\ref{eq:QuantumInterval-valuedProbability-Inclusion}, \yutsung{Check whether
we can explain how to apply
Eq.~\ref{eq:QuantumInterval-valuedProbability-Inclusion} in this situation
better.} we must have
one of the $\bar{\mu}(\proj{\psi_i})=\necess$ and all others mapped to
$\imposs$. Given any state~$\rho$ that is consistent with this QIVPM
on all the projectors in $\Omega$, we have by Eq.~(\ref{eq:hbar}) that
$\mu_{\rho}^{\mathrm{B}}$ must also map exactly one of the projectors
in $\Omega$ to 1 and all others to 0. If an observable has a spectral
decomposition along $\Omega$ then, by
Eq.~(\ref{eq:quantum-expectation}), its expectation value relative to
$\mu_{\rho}^{\mathrm{B}}$ is the eigenvalue $\lambda$ whose projector
is mapped to 1. It therefore follows, by
Eq.~(\ref{eq:quantum-interval-expectation-core}), that the expectation
value relative to the 0-deterministic~${\bar{\mu}}$ is fully
determinate and lies in the interval $[\lambda,\lambda]$.

@@

The quantum probability measure~$\mu_{1}$ maps some projectors to
$0$ or $0.5$, which are replaced by $\left[0,0\right]$ or $\left[0.5,0.5\right]$
to get $\bar{\mu}_{3}$. This is replacing two numbers by the smallest
intervals containing them. In general, any set of the numbers can
be replaced by the interval containing them if we carefully pick a
set of overlapping intervals. For example, consider the following
collection of intervals: 
\begin{eqnarray}
\widehat{\mathscr{I}}_{n} & = & \set{\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\{\imposs,\necess\}
\end{eqnarray}
which are overlapping intervals with length~$1/n$ except the intervals
at length~$\left(2n\right)^{-1}$ at each end. Then, we can define
a QIVPM mapping to $\widehat{\mathscr{I}}_{n}$ based on any Born
rule probability~$\mu_{\rho}^{\mathrm{B}}$ on a three-dimensional
Hilbert space by replacing the probability~$\mu_{\rho}^{\mathrm{B}}\left(P\right)$
with the interval containing it, $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$.
Since intervals in $\widehat{\mathscr{I}}_{n}$ are overlapping, if
$\mu_{\rho}^{\mathrm{B}}\left(P\right)$ are contained in both intervals~$[\ell_{1},r_{1}]$
and $[\ell_{2},r_{2}]$, $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
will be $[\ell_{i},r_{i}]$ if $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
is closer to the center of $[\ell_{i},r_{i}]$. When there is a tie,
we set $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)=\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]$
if $k$ even. For example, consider $n=2$, 
\begin{eqnarray}
\widehat{\mathscr{I}}_{1} & = & \set{\left[\frac{k-1}{4},\frac{k+1}{4}\right]\cap\left[0,1\right]}{k=0,\ldots,4}\nonumber \\
 &  & \cup\{\imposs,\necess\}\nonumber \\
 & = & \left\{ \imposs,\left[0,\frac{1}{4}\right],\left[0,\frac{1}{2}\right],\left[\frac{1}{4},\frac{3}{4}\right],\right.\nonumber \\
 &  & \left.\left[\frac{1}{2},1\right],\left[\frac{1}{4},1\right],\necess\right\} \,.
\end{eqnarray}
$\widehat{\iota}_{2}\circ\mu_{1}$ can be computed by Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute the QIVPM~$\widehat{\iota}_{2}\circ\mu_{1}$
from $\mu_{1}$ on the minimal subspace of events containing $\proj{0}$,
$\proj{1}$, and $\proj{2}$.}
\begin{tabular}{ccc}
\toprule 
$P$ & $\mu_{1}\left(P\right)$ & $\widehat{\iota}_{2}\left(\mu_{1}\left(P\right)\right)$\tabularnewline
\midrule
\midrule 
$\mathbb{0}$ & $0$ & $\left[0,0\right]$\tabularnewline
\midrule 
$\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\proj{2}$ & $0$ & $\left[0,0\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{2}$ & $1$ & $\left[1,1\right]$\tabularnewline
\midrule 
$\mathbb{1}$ & $1$ & $\left[1,1\right]$\tabularnewline
\bottomrule
\end{tabular}
\end{table}
It is clear that $\widehat{\iota}_{2}\circ\mu_{1}$ satisfies Eqs.~(\ref{eq:QIVPM-constraints}).
For Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}),

@@

Although Gleason's theorem does not hold for every QIVPM, we can show
that for a certain class of QIVPMs, the collection of states
consistent with the projectors get closer and closer to each other as
the intervals get narrower and narrower, eventually collapsing to a
single state when the intervals become infinitely precise real
numbers. To state this result, we introduce a few technical notions.

We call a state $\rho$ a \emph{totally mixed state} if
$0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$ for any
$P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $. We also
define a notion of distance based on the following norm of a
matrix~$A$ as
$\left\Vert A\right\Vert
=\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$~\citep{544199}. Now
consider the following collection of intervals:
\begin{eqnarray}
\widehat{\mathscr{I}}_{n} & = & \set{\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\{ \imposs, \necess \}
\end{eqnarray}
which are overlapping intervals with length~$1/n$ except the intervals
at length~$\left(2n\right)^{-1}$ at each end.

A totally mixed state induces a QIVPM over the intervals above as
follows. We first define a map $\widehat{\iota}_{n}$ from the interval
$[0,1]$ to the collection of intervals as:
\begin{itemize}
\item when $x=0$ or $x=1$, $\widehat{\iota}_{n}\left(x\right)=\left[x,x\right]$;
\item when $0<x<1$, $\widehat{\iota}_{n}\left(x\right)=\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]$,
where $k$ is defined by one of the following:
\begin{enumerate}
\item $k$ even and $k-0.5\le2nx\le k+0.5$;
\item $k$ odd and $k-0.5<2nx<k+0.5$.
\end{enumerate}
\end{itemize}

It is now not difficult to verify that for any totally mixed
state~$\rho$ on a three-dimensional Hilbert space,
$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ is a
QIVPM. Furthermore, we have:
\begin{itemize}
\item If $\left\Vert \rho-\rho'\right\Vert \le1/\left(12n\right)$,
$\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$.
\item If $\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
then $\left\Vert \rho-\rho'\right\Vert \le1/n$.
\end{itemize}
In other words, for the states consistent with this QIVPM,
$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ will become closer and
closer to $\rho'$ as $n$ goes to infinity. In the limiting case when
the intervals are infinitely precise, i.e.,
$\set{\left[x,x\right]}{x\in\left[0,1\right]}$, any QIVPM can be
recast as a real-valued one and will have a unique state consistent
with it on every projector when $d\ge3$.

>>>>

\begin{definition}[QIVPM]
  Assume a collection of intervals $\mathscr{I}$ including $\imposs$
  and $\necess$ with addition and scalar multiplication defined as
  follows:
  \begin{subequations}\label{eq:interval-operations}
  \begin{eqnarray}
   &  & [\ell_{1},r_{1}]+[\ell_{2},r_{2}]=[\ell_{1}+\ell_{2},r_{1}+r_{2}]\textrm{ and}\\
   &  & x[\ell,r]=\begin{cases}
  [x\ell,xr] & \textrm{for }x\ge0\,;\\{}
  [xr,x\ell] & \textrm{for }x\le0\,.
  \end{cases}
  \end{eqnarray}
  \end{subequations}
  Then we take a QIVPM~$\bar{\mu}$ to be an assignment of an interval to each
  event (projection operator~$P$) subject to the following constraints:
  \begin{subequations}\label{eq:QIVPM-constraints}
  \begin{eqnarray}
   &  & \bar{\mu}(\mathbb{0})=\imposs\,,\\
   &  & \bar{\mu}(\mathbb{1})=\necess\,,\label{eq:necess}\\
   &  & \bar{\mu}\left(\mathbb{1}-P\right)= \necess -\bar{\mu}\left(P\right)\,,
  \end{eqnarray}
  \end{subequations}
  and satisfying for each pair of \emph{commuting} projectors~$P_0$
  and~$P_1$ with $P_0P_1=P_1P_0$,
\begin{equation}
\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,.
\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{definition}
\noindent The first three constraints,
Eqs.~(\ref{eq:QIVPM-constraints}), are the direct counterpart of the
corresponding ones for classical IVPMs.  With the understanding that
the union of classical sets $E_1\cup E_2$ is replaced by
$P_0+P_1-P_0P_1$ in the case of quantum projection
operators~\cite{Griffiths2003}, the last condition,
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}), is a
direct counterpart of the convexity condition of
Eq.~(\ref{eq:classicalconvex}). Thus our definition of QIVPMs merges
classical IVPMs with quantum probability measures.

@@

The expectation value of an observable relative to a 0-deterministic
QIVPM is fully determinate and is equal to one of the eigenvalues of
that observable. To see this, note that given an orthonormal basis
$\Omega=\{ \ket{\psi_0},\ket{\psi_1},\ldots\}$, a 0-deterministic
QIVPM must map exactly one of the projectors $\proj{\psi_i}$ to
$\necess$ and all others to $\imposs$. This is because by
Eq.~(\ref{eq:necess}), we have
$\bar{\mu}(\Sigma_i{\proj{\psi_i}})=\necess$ and by inductively applying
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}), \yutsung{Check whether
we can explain how to apply
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}) in this situation
better.} we must have
one of the $\bar{\mu}(\proj{\psi_i})=\necess$ and all others mapped to
$\imposs$. Given any state~$\rho$ that is consistent with this QIVPM
on all the projectors in $\Omega$, we have by Eq.~(\ref{eq:hbar}) that
$\mu_{\rho}^{\mathrm{B}}$ must also map exactly one of the projectors
in $\Omega$ to 1 and all others to 0. If an observable has a spectral
decomposition along $\Omega$ then, by
Eq.~(\ref{eq:quantum-expectation}), its expectation value relative to
$\mu_{\rho}^{\mathrm{B}}$ is the eigenvalue $\lambda$ whose projector
is mapped to 1. It therefore follows, by
Eq.~(\ref{eq:quantum-interval-expectation-core}), that the expectation
value relative to the 0-deterministic~${\bar{\mu}}$ is fully
determinate and lies in the interval $[\lambda,\lambda]$.

@@

The quantum probability measure~$\mu_{1}$ maps some projectors to
$0$ or $0.5$, which are replaced by $\left[0,0\right]$ or $\left[0.5,0.5\right]$
to get $\bar{\mu}_{3}$. This is replacing two numbers by the smallest
intervals containing them. In general, any set of the numbers can
be replaced by the interval containing them if we carefully pick a
set of overlapping intervals. For example, consider the following
collection of intervals: 
\begin{eqnarray}
\widehat{\mathscr{I}}_{n} & = & \set{\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\{\imposs,\necess\}
\end{eqnarray}
which are overlapping intervals with length~$1/n$ except the intervals
at length~$\left(2n\right)^{-1}$ at each end. Then, we can define
a QIVPM mapping to $\widehat{\mathscr{I}}_{n}$ based on any Born
rule probability~$\mu_{\rho}^{\mathrm{B}}$ on a three-dimensional
Hilbert space by replacing the probability~$\mu_{\rho}^{\mathrm{B}}\left(P\right)$
with the interval containing it, $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$.
Since intervals in $\widehat{\mathscr{I}}_{n}$ are overlapping, if
$\mu_{\rho}^{\mathrm{B}}\left(P\right)$ are contained in both intervals~$[\ell_{1},r_{1}]$
and $[\ell_{2},r_{2}]$, $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
will be $[\ell_{i},r_{i}]$ if $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
is closer to the center of $[\ell_{i},r_{i}]$. When there is a tie,
we set $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)=\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]$
for even $k$. For example, when $n=2$, 
\begin{eqnarray}
\widehat{\mathscr{I}}_{2} & = & \set{\left[\frac{k-1}{4},\frac{k+1}{4}\right]\cap\left[0,1\right]}{k=0,\ldots,4}\nonumber \\
 &  & \cup\{\imposs,\necess\}\nonumber \\
 & = & \left\{ \imposs,\left[0,\frac{1}{4}\right],\left[0,\frac{1}{2}\right],\left[\frac{1}{4},\frac{3}{4}\right],\right.\nonumber \\
 &  & \left.\left[\frac{1}{2},1\right],\left[\frac{1}{4},1\right],\necess\right\} \,.
\end{eqnarray}
It is straightforward to verify $\widehat{\iota}_{2}\circ\mu_{1}$
is a QIVPM. It will be denoted by $\bar{\mu}_{4}$ whose values on
the minimal subspace of events containing $\proj{0}$, $\proj{1}$,
and $\proj{2}$, $\events_{C}$, is computed by Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\bar{\mu}_{4}$ based on $\mu_{1}$
on $\events_{C}$.}
\begin{tabular}{ccc}
\toprule 
$P$ & $\mu_{1}\left(P\right)$ & $\bar{\mu}_{4}\left(P\right)$\tabularnewline
\midrule
\midrule 
$\mathbb{0}$ & $0$ & $\left[0,0\right]$\tabularnewline
\midrule 
$\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\proj{2}$ & $0$ & $\left[0,0\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{2}$ & $1$ & $\left[1,1\right]$\tabularnewline
\midrule 
$\mathbb{1}$ & $1$ & $\left[1,1\right]$\tabularnewline
\bottomrule
\end{tabular}
\end{table}

The QIVPM~$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ is well-behavior
in a sense that the collection of states consistent it get closer
and closer to $\rho$ as the intervals get narrower and narrower,
eventually collapsing to $\rho$ when the intervals become infinitely
precise real numbers. To state this result, we need some more technical
notions. We first define a notion of distance based on the following
norm of a matrix~$A$ as $\left\Vert A\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$~\citep{544199}.
Then, for any state~$\rho$ satisfying $0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$
for any $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $,
any other state~$\rho'$ consistent with $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$
must close to $\rho$ in a sense that $\left\Vert \rho-\rho'\right\Vert \le1/\left(12n\right)$.
Reversely, if a state~$\rho'$ is close to $\rho$ in a sense that
$\left\Vert \rho-\rho'\right\Vert \le1/n$, then $\rho'$ is consistent
with $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$. In other
words, for the states consistent with this QIVPM, $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho'$ as $n$ goes to infinity.
In the limiting case when the intervals are infinitely precise, i.e.,
$\set{\left[x,x\right]}{x\in\left[0,1\right]}$, any QIVPM can be
recast as a real-valued one and will have a unique state consistent
with it on every projector.

@@
